A \proglang{Python}/\proglang{Zig} optimized and customizable implementation for the $\rho_{DCCA}$ and $DMC_x^2$ methods

### 2024-11-06 paper_v01/article.tex
P_DCCA function explained

With the mandatory steps explained, some very important optional inputs should be addressed. Starting with the $\rho_{DCCA}$ function (dark green node in Figure \ref{fig:dmc_chart}) represents the input \verb"DCCA_of" of the function. This input requires a 2D array of index. For example: if the \verb"input_data" receives a four columns matrix, with index ranging from $0$ to $3$, that is intended to calculate de $DFA$ for all the series but the $\rho_{DCCA}$ between series of index $0$ and $1$ and also for series $2$ and $3$, the \verb"DCCA_of" input should receive the array \verb"[[0,1], [2,3]]". If no value (or the string \verb"'all'") is given, the function will calculate all possible combinations of $DCCA$ calculations between all the series respecting the index values order, as below:

@@

\begin{Code}
for n_index, n  in enumerate(tws): # for each time scale
  # temporaty allocation arrays
  f2dfa_n = np.full(shape=(data.shape[0] - n, data.shape[1]),
  fill_value=np.nan, dtype=data.dtype)
  dcca_n = np.full(shape=(data.shape[0] - n, DCCA_of.shape[0]), 
  fill_value=np.nan, dtype=data.dtype)
  detrended_mat = np.full(shape=(n + 1, data.shape[1]), 
  fill_value=np.nan, dtype=data.dtype)
  for i in range(data.shape[0] - n): # for each box
      detrended_series( # inputs
                        time_steps[i : i + (n + 1)],  # arr_x
                        data[i : i + (n + 1), :],  # mat_y
                         detrended_mat,  # output
                      )
      f2dfa_n[i] = np.power(detrended_mat, 2).mean(axis=0)
      for j, pair in enumerate(DCCA_of): # for each DCCA pair
          dcca_n[i, j] = (detrended_mat[:, pair[0]] * detrended_mat[:, pair[1]]
                         ).mean(axis=0)
  F_DFA_arr[n_index, :] = np.sqrt(f2dfa_n.mean(axis=0))
  DCCA_arr[n_index, :] = dcca_n.mean(axis=0)
  # calculation of P_DCCA
  P_DCCA_output_function(n_index, DCCA_of, F_DFA_arr, DCCA_arr, P_DCCA_arr)
\end{Code}

The first \verb"for" loop in the code operates over the values of the \verb"tws" input, asserting that the step 6 of the $DFA$ and $DCCA$ methods in Section \ref{sec:calculations}. Three temporary arrays are allocated for each time scale, \verb"f2dfa_n"

>>>>

With the mandatory steps explained, some very important optional inputs should be addressed. Starting with the $\rho_{DCCA}$ function (dark green node in Figure \ref{fig:dmc_chart}) represents the input \verb"DCCA_of" of the function. This input requires a 2D array of integers, each row is a pair of index, related to the \verb"data_input" matrix. For example: if the \verb"input_data" receives a four columns matrix, with index ranging from $0$ to $3$, that is intended to calculate de $DFA$ for all the series but the $\rho_{DCCA}$ between series of index $0$ and $1$ and also for series $2$ and $3$, the \verb"DCCA_of" input should receive the array \verb"[[0,1], [2,3]]". If no value (or the string \verb"'all'") is given, the function will calculate all possible combinations of $DCCA$ calculations between all the series respecting the index values order, as below:

@@

\begin{Code}
for n_index, n  in enumerate(tws): # for each time scale
  # temporaty allocation arrays
  f2dfa_n = np.full(shape=(data.shape[0] - n, data.shape[1]),
  fill_value=np.nan, dtype=data.dtype)
  dcca_n = np.full(shape=(data.shape[0] - n, DCCA_of.shape[0]), 
  fill_value=np.nan, dtype=data.dtype)
  detrended_mat = np.full(shape=(n + 1, data.shape[1]), 
  fill_value=np.nan, dtype=data.dtype)
  for i in range(data.shape[0] - n): # for each box
      detrended_series( # inputs
                        time_steps[i : i + (n + 1)],  # arr_x
                        data[i : i + (n + 1), :],  # mat_y
                         detrended_mat,  # output
                      )
      f2dfa_n[i] = np.power(detrended_mat, 2).mean(axis=0)
      for j, pair in enumerate(DCCA_of): # for each DCCA pair
          dcca_n[i, j] = (detrended_mat[:, pair[0]] * detrended_mat[:, pair[1]]
                         ).mean(axis=0)
  F_DFA_arr[n_index, :] = np.sqrt(f2dfa_n.mean(axis=0))
  DCCA_arr[n_index, :] = dcca_n.mean(axis=0)
  # calculation of P_DCCA
  P_DCCA_output_function(n_index, DCCA_of, F_DFA_arr, DCCA_arr, # Inputs
  P_DCCA_arr) # Output
\end{Code}

The first \verb"for" loop in the code operates over the values of the \verb"tws" input, asserting that the step 6 of the $DFA$ and $DCCA$ methods in Section \ref{sec:calculations}. Three temporary arrays are allocated for each time scale. The first, \verb"f2dfa_n", to store the calculations of the step 4 of the $DFA$. The number of lines of this array correspond to the number of boxes in the current time scale ($N - n$) and the columns is the number of series in the analysis. Array  \verb"dcca_n", holds the values calculated in the step 4 of the $DCCA$. The number of lines also correspond to the number of boxes but the number of columns equals the number of pairs (rows) in the \verb"DCCA_of" input. The \verb"detrended_mat" array has a number of lines is the number of points in a time scale box ($n +1$) and column count also equal to the number of time series. This last array is very important to the implementation goals.

In each box, the \verb"detrended_series" function execute a one degree polynomial fit, subtract the values of the serie with the value given by the interpolated curve and stores this values in the \verb"detrended_mat". After the $DFA$ is calculated. With all the detrended values calculated, another loop, for all the rows in the \verb"DCCA_of", uses the \verb"detrended_mat" and the $DFA$ arrays to calculate the $DCCA$. This approach avoid redundant calculation. If the \verb"DCCA_of" is $[[0,1], [0,2], [1,2]]$, the detrended and $DFA$ values for each serie are calculated just once. If in a set of four series, a research intend to calculate the $DCCA$ between the first serie and all the others along with the two last series, inputting  \verb"DCCA_of" $= [[0,1], [0,2], [0,3], [2,3]]$. In this case, the \verb"detrended_mat" will be calculated just once for each series and the $DCCA$ between series $[[1,2], [1,3]]$ will be skipped, revealing a very customizable algorithm.

The $\rho_{DCCA}$ function, named \verb"P_DCCA_output_function" in the code is, in fact a pointer to other functions. One that outputs the  $\rho_{DCCA}$ results in the form of a table (rows for each time scale and columns for each \verb"DCCA_of" pair) and the other outputs it the form of a 3D matrix, where each level is the matrix in Equation \ref{eq:p_dcca_matrix} for one of the time scales. This behavior is driven by the \verb"P_DCCA_output_matrix" input (represented as the light green node in Figure \ref{fig:pdcca_chart}), where \verb"False" means table output and \verb"True" matrix output. This is very convenient for calculating the $DMC_x^2$.

The $DMC_x^2$ function runs the $\rho_{DCCA}$ with \verb"P_DCCA_output_matrix" set to \verb"True".
